---
title: "Week 4 Project - Getting and Cleaning"
author: "J. Rafael Zárate-Salazar"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

# Load packages

```{r }
library(tidyverse)
library(here)
```

# Load Data Base

# Get Test Data

```{r}

xtest <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","test","X_test.txt")
                     ,header = FALSE)

ytest <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","test","y_test.txt")
                      ,header = FALSE)

```

# Get Train Data

```{r}

xtrain <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","train","X_train.txt")
                     ,header = FALSE)

ytrain <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","train","y_train.txt")
                      ,header = FALSE)

```

# Get Features Data

```{r}

features <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","features.txt")
                       ,header = FALSE)

```

# Get Activity Data

```{r}

activity <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","activity_labels.txt")
                       ,header = FALSE)

```

# Get Subject Data

## Substest

```{r}

subtest <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","test","subject_test.txt")
                      ,header = FALSE)

subtest <- subtest %>%
  dplyr::rename(subjectID = V1)

```

## Substrain

```{r}

subtrain <- read.table( file = here::here("Getdata_projectfile","UCI_HAR_Dataset","train","subject_train.txt")
                     ,header = FALSE)

subtrain <- subtrain %>%
  dplyr::rename(subjectID = V1)

```

# Add column names to both train and test data

```{r}

features <- features[,2]
featrasp <- t(features)

colnames(xtrain) <- featrasp
colnames(xtest) <- featrasp

```

# Rename activity columns to id and actions

```{r}

colnames(activity) <- c('id','actions')

```

# Row bind xtrain and xtest 

```{r}

combineX <- rbind(xtrain, xtest) 

```

# Row bind ytrain and ytest

```{r}

combineY <- rbind(ytrain, ytest)

```

# Row bind subject train and subject test

```{r}

combineSubj <- rbind(subtrain,subtest)

```

# Column bind Y and X, except for activity 

```{r}

joinXY <- cbind(combineY,combineX, combineSubj)

```

# 1. Joining the above data frame with the activity

```{r, message=FALSE,warning=FALSE}

db <- merge(joinXY, activity, by.x = 'V1', by.y = 'id')

```


# 2. Getting the mean and standard deviation

```{r}

colNames <- colnames(db)

db.final <- db %>%
  dplyr::select(actions, subjectID, stringr::str_subset(colNames,"\\bmean\\b|\\bstd\\b"))

```

# Transform activity to a factor variable 

```{r}

db.final$actions <- as.factor(db.final$actions)

```

# 3 and 4. Use descriptive activity names to name the activities in the data set

```{r}

colnames(db.final) <- gsub("^t", "time", colnames(db.final))
colnames(db.final) <- gsub("^f", "frequency", colnames(db.final))
colnames(db.final) <- gsub("Acc", "Accelerometer", colnames(db.final))
colnames(db.final) <- gsub("Gyro", "Gyroscope", colnames(db.final))
colnames(db.final) <- gsub("Mag", "Magnitude", colnames(db.final))
colnames(db.final) <- gsub("BodyBody", "Body", colnames(db.final))

```

# 5. Creates a second data set with the average of each variable for activity and subject.

```{r}

db.final <- db.final %>% 
  dplyr::group_by(actions,subjectID) %>% 
  dplyr::arrange(desc(actions)) %>% 
  dplyr::summarise_all(mean) %>% 
  dplyr::select(subjectID,everything())

db.final

```

# Save binary data

```{r}

# save(db.final, file = here::here("Getdata_projectfile","UCI_HAR_Dataset","tidydata.rda"))

```


# Text file for final output

```{r}

# write.table(db.final, file = here::here("Getdata_projectfile","UCI_HAR_Dataset","tidydata.txt")
#                                     ,row.name = FALSE)

```
